### 2023-03-11 scripts/01-data_import.qmd

---
title: "Import data from the Toronto Paramedic Service via Open Data Toronto"
format: html
author: "Sebastian Rodriguez"
---
```{r}
#### Preamble ####
# Purpose: Import data from the Toronto Paramedic Service via Open Data Toronto
# Author: Sebastian Rodriguez
# Email: me@srod.ca
# Date: 3 February 2023
# GitHub: https://github.com/seb646/toronto-paramedic-responses

#### Workspace set-up ####
library(opendatatoronto)
library(tidyverse)
library(dplyr)
```

```{r}
#| warning: false
#### Fetch "Land Ambulance Response Time Standard" from Open Data Toronto
# Link: https://open.toronto.ca/dataset/land-ambulance-response-time-standard 
raw_response_data <-
  list_package_resources("8b068db1-95b8-4ea9-b942-8a7ce3a5fbd6") |>
  filter(name == "Response Time (2013 - 2021)") |>
  get_resource()

# Write data to a .csv file
write_csv(
  x = raw_response_data$`Land Ambulance Services`,
  file = "../inputs/data/raw_response_data.csv"
)
```

```{r}
#| warning: false
#### Fetch "Pre-Hospital Emergency Care Performance Metrics" from Open Data Toronto ####
# Link: https://open.toronto.ca/dataset/pre-hospital-emergency-care-performance-metrics
raw_performance_data <-
  list_package_resources("174ace54-bcc7-4885-85c1-76094ab4a0ca") |>
  filter(name == "Pre-hospital emergency care performance metrics 2014-2021") |>
  get_resource() |>
  filter(`Pre-Hospital Emergency Care Type` == "Emergency Request for Service")

# Write data to a .csv file
write_csv(
  x = raw_performance_data,
  file = "../inputs/data/raw_performance_data.csv"
)
```

```{r}
#| warning: false
#### Fetch "Paramedic Services Incident Data" from Open Data Toronto ####
# Link: https://open.toronto.ca/dataset/paramedic-services-incident-data
raw_incident_data <-
  list_package_resources("c21f3bd1-e016-4469-abf5-c58bb8e8b5ce") |>
  filter(name == "paramedic-services-incident-data-2017-2021") |>
  get_resource()

# Write data to a .csv file
write_csv(
  x = raw_incident_data$`2021`,
  file = "../inputs/data/raw_incident_data.csv",
)

# Note: creating this .csv may take some time as it's a large data set
```