### 2023-03-11 scripts/02-data_cleaning.qmd

---
title: "Import raw csv data, clean it, and generate new csv files"
format: html
author: "Sebastian Rodriguez"
---
```{r}
#### Preamble ####
# Purpose: Import raw csv data, clean it, and generate new csv files
# Author: Sebastian Rodriguez
# Email: me@srod.ca
# Date: 3 February 2023
# GitHub: https://github.com/seb646/toronto-paramedic-responses

#### Workspace set-up ####
library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(janitor)
```

```{r}
#### Fetch and clean the raw response data ####
raw_happy_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, happy) |>
  filter(happy >= 2) |> 
  count(year, happy) |>
  group_by(year) |>
  summarise(people = sum(n))

raw_directions_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, directns) |>
  filter(directns >= 1, directns <= 4) |> 
  count(year, directns) |>
  group_by(year) |>
  summarise(people = sum(n))

raw_give_homeless_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, givhmlss) |>
  filter(givhmlss >= 1, givhmlss <= 4) |>
  count(year, givhmlss) |>
  group_by(year) |>
  summarise(people = sum(n))

colors <- c(
  "Happiness" = "steelblue", 
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen")

ggplot() +
  # Define axis labels
  labs(
    y = "People", 
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = raw_happy_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = raw_directions_data, aes(x = year, y = people, color = "Give Directions to Stranger")) +
  geom_line(data = raw_give_homeless_data, aes(x = year, y = people, color = "Give Money to Homeless")) +
  scale_x_continuous(breaks=c(2002,2004,2012,2014)) +
  scale_color_manual(values = colors)

```

```{r}
#### Fetch and clean the raw response data ####
raw_happy_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, happy) |>
  filter(happy == 3) |> 
  count(year, happy) |>
  group_by(year) |>
  summarise(people = sum(n))

raw_directions_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, directns) |>
  filter(directns >= 5) |> 
  count(year, directns) |>
  group_by(year) |>
  summarise(people = sum(n))

raw_give_homeless_data <- readr::read_csv(here::here("inputs/data/raw_data.csv"), show_col_types = FALSE) |>
  # Change column naming format
  clean_names() |> 
  select(year, givhmlss) |>
  filter(givhmlss >= 5) |>
  count(year, givhmlss) |>
  group_by(year) |>
  summarise(people = sum(n))

colors <- c(
  "Happiness" = "steelblue", 
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen")

ggplot() +
  # Define axis labels
  labs(
    y = "People", 
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = raw_happy_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = raw_directions_data, aes(x = year, y = people, color = "Give Directions to Stranger")) +
  geom_line(data = raw_give_homeless_data, aes(x = year, y = people, color = "Give Money to Homeless")) +
  scale_x_continuous(breaks=c(2002,2004,2012,2014)) +
  scale_color_manual(values = colors)

```