### 2023-03-15 outputs/paper/paper.qmd
Updated paper graphs

---
title: "Does Happiness Lead to Altruism: A review of US General Social Survey data"
format: pdf
author: "Sebastian Rodriguez, Iz Leitch, Randall Ni"
abstract: "Something about happiness and altruism" 
date: today
date-format: long
editor: visual
toc: TRUE
number-sections: TRUE
bibliography: references.bib
thanks: "Code and data supporting this analysis is available at: https://github.com/seb646/happiness-and-altruism"
---

```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Preamble ####
# Title: Does Happiness Lead to Altruism: A review of US General Social Survey data
# Author: Sebastian Rodriguez, Iz Leitch, Randall Ni
# Email: me@srod.ca
# Date: 3 February 2023
# GitHub: https://github.com/seb646/happiness-and-altruism

#### Workspace set-up ####
# Define required libraries 
library(lubridate)
library(tidyverse)
library(dplyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(readr)

# Fetch the the clean response data
happy_positive_data <- readr::read_csv(here::here("inputs/data/clean/happy_positive_data.csv"), show_col_types = FALSE)

# Fetch clean performance data
directions_positive_data <- readr::read_csv(here::here("inputs/data/clean/directions_positive_data.csv"), show_col_types = FALSE)

# Fetch the clean incident response
homeless_positive_data <- readr::read_csv(here::here("inputs/data/clean/homeless_positive_data.csv"), show_col_types = FALSE)

# Fetch the the clean response data
happy_negative_data <- readr::read_csv(here::here("inputs/data/clean/happy_negative_data.csv"), show_col_types = FALSE)

# Fetch clean performance data
directions_negative_data <- readr::read_csv(here::here("inputs/data/clean/directions_negative_data.csv"), show_col_types = FALSE)

# Fetch the clean incident response
homeless_negative_data <- readr::read_csv(here::here("inputs/data/clean/homeless_negative_data.csv"), show_col_types = FALSE)
```

## Introduction

\newpage

## General Social Survey Data

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Happiness vs Giving Directions to Strangers"
#| label: fig-1

colors <- c(
  "Happiness" = "steelblue",
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen"
)

ggplot() +
  # Define axis labels
  labs(
    y = "People",
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = happy_positive_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = directions_positive_data, aes(x = year, y = people, color = "Give Directions to Stranger")) +
  scale_x_continuous(breaks = c(2002, 2004, 2012, 2014)) +
  scale_color_manual(values = colors)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Happiness vs Giving Money to Homeless"
#| label: fig-2

colors <- c(
  "Happiness" = "steelblue",
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen"
)

ggplot() +
  # Define axis labels
  labs(
    y = "People",
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = happy_positive_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = homeless_positive_data, aes(x = year, y = people, color = "Give Money to Homeless")) +
  scale_x_continuous(breaks = c(2002, 2004, 2012, 2014)) +
  scale_color_manual(values = colors)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Unhappiness vs. Not Giving Directions to Strangers"
#| label: fig-3

colors <- c(
  "Happiness" = "steelblue",
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen"
)

ggplot() +
  # Define axis labels
  labs(
    y = "People",
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = happy_negative_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = directions_negative_data, aes(x = year, y = people, color = "Give Directions to Stranger")) +
  scale_x_continuous(breaks = c(2002, 2004, 2012, 2014)) +
  scale_color_manual(values = colors)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Unhappiness vs. Not Money to Homeless"
#| label: fig-4

colors <- c(
  "Happiness" = "steelblue",
  "Give Directions to Stranger" = "darkred",
  "Give Money to Homeless" = "darkgreen"
)

ggplot() +
  # Define axis labels
  labs(
    y = "People",
    x = "Year",
    color = "Legend"
  ) +

  # Set the theme
  theme_minimal() +

  geom_line(data = happy_negative_data, aes(x = year, y = people, color = "Happiness")) +
  geom_line(data = homeless_negative_data, aes(x = year, y = people, color = "Give Money to Homeless")) +
  scale_x_continuous(breaks = c(2002, 2004, 2012, 2014)) +
  scale_color_manual(values = colors)
```
## Discussion

\newpage

## Inclusive Acknowledgements

**Land Acknowledgement**

We wish to acknowledge the land on which this data was collected and analysed. For thousands of years, it has been the traditional land of the Huron-Wendat, the Seneca, and the Mississaugas of the Credit. Today, this meeting place is still the home to many Indigenous people from across Turtle Island and we are grateful to have the opportunity to work on this land.

This statement was originally prepared in consultation with First Nations House and the Elders Circle for use at the University of Toronto [@cite_land_acknowledgement].\

**Data Acknowledgement**

The data used in this paper was obtained through the General Social Survey (GSS), a project of the independent research organization NORC at the University of Chicago, with principal funding from the National Science Foundation. This paper uses three data sets from the NORC's GSS:

-   General Happiness[^1]
-   Has Given Directions to a Stranger[^2]
-   Has Given Money or Food to a Homeless Person[^3]

[^1]: General Happiness data is available at: https://gssdataexplorer.norc.org/variables/434/vshow

[^2]: Has Given Directions to a Stranger data is available at: https://gssdataexplorer.norc.org/variables/2886/vshow

[^3]: Has Given Money or Food to a Homeless Person data is available at: https://gssdataexplorer.norc.org/variables/2878/vshow

**Resources Acknowledgment**

The primary tool used to analyse data in this paper is R, an open-source statistical programming language [@cite_tool_r]. The paper also uses a number of R packages, including: dplyr [@cite_tool_dplyr], ggplot2 [@cite_tool_ggplot2], here [@cite_tool_here], janitor [@cite_tool_janitor], kableExtra [@cite_tool_kableextra], knitr [@cite_tool_knitr], lubridate [@cite_tool_lubridate], opendatatoronto [@cite_tool_opendatatoronto], readr [@cite_tool_readr], RColorBrewer [@cite_tool_rcolorbrewer], scales [@cite_tool_scales], and tidyverse [@cite_tool_tidyverse]. \newpage

## References

<!-- Comment required to show references as a numbered section -->